---
title: "linear.regression (by Shashi and Rojan)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(linear.regression)
```





### Group members

* Shashi Nagarajan Iyer
* Rojan Karakaya


# Usage

The [linear.regression] package is an attempt at a **limited** reconstruction of the [lm] package.

We define a model using the same notation that we would use for [lm]

```{r}
data("iris")
linr = linreg(Petal.Length ~ Species + Sepal.Width, data = iris)
```

For a quick overview of our model (both the function call and the estimated coefficients) we can use the print(*model*) command:

```{r}
print(linr)
```

If we only want a quick glance at the estimated coefficients, coef(*model*) works just as well:

```{r}
coef(linr)
```

For a more in-depth summary of our model we use the summary(*model*) function:

```{r}
summary(linr)
```

we can get our predicted (i.e. fitted) values with pred(*model*)

```{r}
head(pred(linr), n=3)
```

plot(*model*) returns the "Residuals vs Fitted" as well as the "Scale-Location" plots:

```{r}
plot(linr)
```

